---
title: "p8105_hw2_az2356"
author: "Amelia Zhao"
date: "9/19/2019"
output: github_document
---
# This is my markdown file for Homework 2.

# Problem 1

Setting up the necessary packages.
```{r}

library(readxl)
library(tidyverse)

```

Working on the Mr. Trash Wheel data:

- Creating a relative directory

- Identifying the Mr. Trash Wheel sheet within the excel file

- Removing rows without dumpster-specific data (totals) and tidying data

```{r}

"./data/trashwheel.xlsx"

trashwheel <- 
  read_excel("data/trashwheel.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names(dat = .) %>%
  rename(.data = ., volume = volume_cubic_yards, styrofoam = polystyrene, weight = weight_tons) %>%
  drop_na(dumpster)
  mutate(
    trashwheel, sports_balls = as.integer(sports_balls)
    )

```



Working on the 2017 and 2018 Precipitation data:

- Creating a year variable, omitting rows without precipitation data

- Combining the 2017 and 2018 data, converting month from a numeric to a character variable with the month name

```{r}

rain2017 <- 
  read_excel("data/trashwheel.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names(dat = .) %>% 
  rename(.data = ., inchesrain = total) %>% 
  mutate(year = 2017)


rain2018 <-
  read_excel("data/trashwheel.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names(dat = .) %>% 
  rename(.data = ., inchesrain = total) %>% 
  mutate(year = 2018)


rain2yr = 
  bind_rows(rain2017, rain2018) %>% 
  mutate(month = month.name[month])

```

Mr. Trash Wheel has been hard at work! In 5 years (May 2014 to June 2019), Mr. Trash Wheel has collected `r sum(trashwheel$weight)` tons, or `r sum(trashwheel$volume)` cubic yards, of trash from the Baltimore Inner Harbor! That is a ton of trash! Mr. Trash Wheel has filled `r nrow(trashwheel)` dumpsters! From the trash collected and put into Mr. Trash Wheel's dumpster, the median number of sports balls was `r median(trashwheel$sports_balls)` in 2017, and the average number of chip bags was `r mean(trashwheel$chip_bags)`. That's a lot of chips. 

From the `r nrow(rain2yr)` observations of precipitation data collected, we can see that there were `r sum(rain2018$inchesrain)` inches of rain in 2018, compared to `r sum(rain2017$inchesrain)` in 2017. 2018 was a wet year!




# Problem 2

Cleaning the datasets: 

- Standardizing year and month across datasets

- Creating "president" variable with "gop", or "dem"

- Merging the politician, S&P, and unemployment data sets

```{r}
## Pols data: the number of politicians who are Republican or Democrat at any given month

"./data/538data/pols-month.csv"
polsmonth = 
  read_csv("data/538data/pols-month.csv") %>% 
  separate(data = ., mon, c("year", "month", "day"), "-", convert = TRUE) %>% 
  mutate(
    month = month.name[month],
    president = case_when(
      prez_gop == 1 ~ "gop",
      prez_dem == 1 ~ "dem"
      )
    ) %>% 
  arrange(year) %>% 
  select(-prez_dem, -prez_gop, -day)


## S & P stock index: closing values at the beginning of each month

"./data/538data/snp.csv"
snp =
  read_csv("data/538data/snp.csv") %>% 
  separate(data = ., date, c("month", "day", "year"), "/", convert = TRUE) %>% 
  select(-day) %>% 
  arrange(month) %>% 
  arrange(year) %>% 
  mutate(
    month = month.name[month],
    ) %>% 
  select(year, month, everything()) 
  

## Unemployment data: unemployment percentage in each month of each year
  
"./data/538data/unemployment.csv"
unemploy <-  
  read_csv("data/538data/unemployment.csv") %>% 
  janitor::clean_names(dat = .) %>% 
  pivot_longer(
    data =, 
    jan:dec,
    names_to = "month",
    values_to = "percent"
  )
  


## Merging the datasets
snppols = 
  left_join(polsmonth, snp)


alldata = 
  left_join(snppols, unemploy)

  


```
The final dataset is an aggregate of the 3 datasets all about America: the polsmonth dataset, which contained information about politicians and their party affiliation at each month, the snp dataset, which contained information about the Standard & Poor closing index each month, and the unemployment dataset, which contained the percentage of unemployment each month. 

All of the observations were kept. However, S&P is missing for many timepoints, as this dataset was the smallest, with only `r nrow(snp)` months of data from `r min(snp$year)` to `r max(snp$year)`, compared to the politician data, which had `r nrow(polsmonth)` months of data, from `r min(polsmonth$year)` to `r max(polsmonth$year)`. Thus, the resulting dataset contains `r nrow(alldata)` months of data, with comprehensive data for the politicians, and missing unemployment and S&P data for some months. 



# Problem 3

```{r}
## Baby names data: popularity of baby names.
"./data/babynames.csv"

babynames <- 
  read_csv("data/babynames.csv") %>% 
  janitor::clean_names(dat = .) %>%
  rename(.data = ., name = childs_first_name) %>%
  distinct(.data = .) %>% 
  arrange(year_of_birth) %>% 
  mutate(
    ethnicity = case_when(
      ethnicity == "ASIAN AND PACI" ~ "Asian and Pacific Islander",
      ethnicity == "ASIAN AND PACIFIC ISLANDER" ~ "Asian and Pacific Islander",
      ethnicity == "BLACK NON HISP" ~ "Black Non Hispanic",
      ethnicity == "BLACK NON HISPANIC" ~ "Black Non Hispanic",
      ethnicity == "HISPANIC" ~ "Hispanic",
      ethnicity == "WHITE NON HISP" ~ "White Non Hispanic",
      ethnicity == "WHITE NON HISPANIC" ~ "White Non Hispanic"
      ),
    gender = str_to_title(gender),
    name = str_to_title(name)
  ) 

babynames %>% 
filter(name == "Olivia") %>%
  select(-count, -gender) %>%
  pivot_wider(names_from = year_of_birth,
              values_from = rank)

babynames %>%
filter(gender == "Male", rank == 1) %>%
  select(-count) %>%
  pivot_wider(names_from = year_of_birth,
              values_from = rank)

nameplot = filter(babynames, gender == "Male", ethnicity == "White Non Hispanic", year_of_birth == "2016")
ggplot(nameplot, aes(x = rank, y = count)) + geom_point(colour = "blue") 

```




