---
title: "p8105_hw2_az2356"
author: "Amelia Zhao"
date: "9/19/2019"
output: github_document
---
# This is my markdown file for Homework 2.

# Problem 1

Setting up the necessary packages.
```{r}
library(readxl)
library(tidyverse)
```

Working on the Mr. Trash Wheel data:

- Creating a relative directory

- Identifying the Mr. Trash Wheel sheet within the excel file

- Removing rows without dumpster-specific data (totals) and tidying data
```{r}

"./data/trashwheel.xlsx"

trashwheel <- 
  read_excel("data/trashwheel.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names(dat = .) %>%
  rename(.data = ., volume = volume_cubic_yards, styrofoam = polystyrene, weight = weight_tons) %>%
  drop_na(dumpster)
  mutate(
    trashwheel, sports_balls = as.integer(sports_balls)
    )

```



Working on the 2017 and 2018 Precipitation data:

- Creating a year variable, omitting rows without precipitation data

- Combining the 2017 and 2018 data, converting month from a numeric to a character variable with the month name

```{r}

rain2017 <- 
  read_excel("data/trashwheel.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names(dat = .) %>% 
  rename(.data = ., inchesrain = total) %>% 
  mutate(year = 2017)

rain2018 <-
  read_excel("data/trashwheel.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names(dat = .) %>% 
  rename(.data = ., inchesrain = total) %>% 
  mutate(year = 2018)

rain2yr = 
  bind_rows(rain2017, rain2018) %>% 
  mutate(month = month.name[month])

```

Mr. Trash Wheel has been hard at work! In 5 years (May 2014 to June 2019), Mr. Trash Wheel has collected `r sum(trashwheel$weight)` pounds of trash from the Baltimore Inner Harbor! From the trash collected and put into Mr. Trash Wheel's dumpster, the median number of sports balls was `r median(trashwheel$sports_balls)`, and the average number of chip bags was `r mean(trashwheel$chip_bags)`. That's a lot of chips. 

From the precipitation data collected, we can see that there were `r sum(rain2018$inchesrain)` inches of rain in 2018, compared to `r sum(rain2017$inchesrain)` in 2017. 2018 was a wet year!




# Problem 2
```{r}
## Pols data: the number of politicians who are Republican or Democrat at any given month
"./data/538data/pols-month.csv"
polsmonth = read_csv("data/538data/pols-month.csv")

## Unemployment data: unemployment percentage in each month of each year
"./data/538data/unemployment.csv"
unemploy = read_csv("data/538data/unemployment.csv")

## S & P stock index: closing values at the beginning of each month
"./data/538data/snp.csv"
snp = read_csv("data/538data/snp.csv")


```



# Problem 3

```{r}
## Baby names data: popularity of baby names.
"./data/babynames.csv"
babynames = read_csv(file = "data/babynames.csv")


```




